---
title: "NMDS Rmd"
output: html_document
editor_options: 
  chunk_output_type: console
---
#ls was here
#NMDS of all species together
LP edits: 
- Removed unnecessary code
- Streamlined # of different datasets (shorebird1 vs shorebird2 etc. condensed to be shorebird)
```{r setup, include=FALSE}

# no need to set wd if in a project on your computer cloned from github
# i have changed the file extensions to work with the project and should work across computers

#setwd("~/Desktop/CNP_internship/LS_Data")

##### packages #####
library(readxl)
library(grDevices)
library(tidyverse)
library(patchwork)
library(janitor)
library(ggdark)
library(Hmisc)
library(broom)
library(viridis)
library(lubridate)
library(vegan) #for RDA
library(ggpubr)
library(rstatix)
library(broom)
library(purrr)
library(labdsv) #hellinger transformation

##### load data #####

#shorebird data
shorebird<-read_excel("data/Shorebird_People_Data/Shorebird_People_Data.xlsx")

# target data (needs alignment - spp lists differ by year)
target <- read_excel('data/RI_Long_Term_Data/Photoplot_summary_by_plot_20210414.xlsx')

# alignment info for target data 
align <- read_csv('data/TGT_all.csv')
# see RI_Species_Alighment_Feb21.R for methods
# summary: took species lists for 1990 and 2000, applied to all unique spp.

```

```{r}
##### presets #####
# lists of commonly filtered-for items
cabrsites <- c('CAB1', 'CAB2', 'CAB3') 
zonelist <- c('CHT', unique(target$Zone)[1:4])
zonenames <- c('Balanus/Chthamalus', 'Tetraclita rubescens', 'Mussel', 'Silvetia compressa', 'Pollicipes polymerus')
targetlist <- c('CHTBAL', 'TETRUB', 'MUSSEL', 'SILCOM', 'POLPOL')

# theme
lltheme_dark <- dark_theme_bw() + theme(text = element_text(size = 20),
                              # add more space between panels
                              panel.spacing = unit(1, 'lines'),
                              # no background to wrap panels
                              strip.background = element_blank(),
                              strip.text = element_text(size = 20, hjust = 0),
                              # panel labels outside x axis labels
                              strip.placement = 'outside',
                              # adjust x axis labels
                              axis.text.y = element_text(size = 20),
                              axis.text.x = element_text(size = 20, angle = 45, hjust = 1))

lltheme_light <- theme_bw() + theme(text = element_text(size = 12),
                              # add more space between panels
                              panel.spacing = unit(1, 'lines'),
                              # no background to wrap panels
                              strip.background = element_blank(),
                              strip.text = element_text(size = 12, hjust = 0),
                              # panel labels outside x axis labels
                              strip.placement = 'outside',
                              # adjust x axis labels
                              axis.text.y = element_text(size = 12),
                              axis.text.x = element_text(size = 12, angle = 45, hjust = 1))
```

```{r}
######## tidy: shorebird data #####

#add column with just years
shorebird$Year=format(shorebird$SurveyDate, "%Y")

#filter for just people
shorebird<- ungroup(shorebird) %>% 
  mutate(SurveyYear=year(SurveyDate))%>%
  filter(DataType=="People")%>%
  group_by(SurveyYear, ZoneClass) %>% #separate by year and zone
  summarise(people=mean(DataCount)) %>% #get average counts
  mutate(ZoneName= paste('Zone', ZoneClass),
         Panel = if_else(ZoneName == 'Zone I', 'A', 
                         if_else(ZoneName == 'Zone II', 'B', 'C'))) %>%
  mutate(ZoneName = paste('(', Panel, ') ', ZoneName, sep = '')) %>%
  select(-c(Panel))%>%
  drop_na()
```

```{r}
##### tidy: photoplot #####
# adapted from B. Hong's 2019-2021 Annual Report Code

# keep target raw, retains plot-level data, and tidy
target <- 
  # add generic codes to this set, along with formatted scientific names
  left_join(target, align, by = 'SpeciesCode') %>%
  # only get cabr sites
  filter(SiteCode %in% cabrsites) %>%
  # get rid of old scientific name (align has better one)
  select(-c(Scientific_name.x)) %>%
  # rename scientific name column
  rename(Scientific_name = Scientific_name.y) %>%
  # make plot number a numeric column
  mutate(Plot_num = parse_number(substr(Plot_code, 6,6))) %>%
  # add zone/site column
  mutate(ZoneName = paste('Zone', substr(SiteName, 10, 12)),
         Panel = if_else(SiteName == 'Cabrillo I', 'A', 
                         if_else(SiteName == 'Cabrillo II', 'B', 'C'))) %>%
  mutate(ZoneName = paste('(', Panel, ') ', ZoneName, sep='')) %>%
  # only get unique rows (removes duplicates)
  distinct()

unique(target$ZoneName)==unique(shorebird$ZoneName)

#### group by categories in 1990 ####

# get sum of number of points for each species
target_1990 <- target %>%
  group_by(SiteCode, SiteName, ZoneName, SurveyYear, Zone, Plot_num, Plot_code, Code_1990) %>%
  summarise(cover = sum(N))

# get number of points/quadrat to calculate% cover
target_summary <- target %>%
  group_by(SiteCode, SiteName, ZoneName, SurveyYear, Zone, Plot_num, Plot_code) %>% 
  summarise(points = sum(N))


target_1990 <- left_join(target_1990, target_summary, by = c("SiteCode", "SiteName", "ZoneName", "SurveyYear", "Zone", "Plot_num", "Plot_code")) %>%
  group_by(SiteCode, ZoneName, SiteName, SurveyYear, Zone, Plot_num, Plot_code, Code_1990) %>%
  mutate(pct_cover = (cover/points)*100) %>%
  select(SiteCode, SiteName, ZoneName, SurveyYear, Zone, Plot_num, Plot_code, Code_1990, pct_cover) %>%
  rename(SpeciesCode = Code_1990) %>%
  filter(SiteName %in% c("Cabrillo I", "Cabrillo II", "Cabrillo III"))

# join spp names to 1990 data
spplist <- align %>%
  select(SpeciesCode, Scientific_name) %>%
  distinct()

target_1990 <- left_join(target_1990, spplist, by = "SpeciesCode")

target_1990_1<-target_1990[,-c(6:8)]

# remove extras
remove(target_summary, align, spplist, target)

# get cover of species in their "home" plots
# for line plots and trend lines of home spp over time and vs people
target_90_home <- ungroup(target_1990) %>%
filter(SiteCode %in% cabrsites) %>%
# filter to only tgt spp in their own plots
filter((Zone == 'CHT' & SpeciesCode == 'CHTBAL') |
(Zone == 'MYT' & SpeciesCode == 'MUSSEL') |
(Zone == 'SIL' & SpeciesCode == 'SILCOM') |
(Zone == 'POL' & SpeciesCode == 'POLPOL')) %>%
# get mean + SE cover for each yr
group_by(ZoneName, SurveyYear, Zone, SpeciesCode, Scientific_name) %>%
summarise(Cover_mean = mean(pct_cover), #keep if want each year to have 1 point
Cover_SE = sd(pct_cover)/sqrt(length(pct_cover)))

#rotate from long to wide
target_new <- select(target_1990, -Scientific_name) %>% pivot_wider(names_from = SpeciesCode, values_from = pct_cover)%>%
  mutate_all(~replace(., is.na(.), 0)) %>%
  # remove blank "na" column
  select(-`NA`)

#combine shorebird data and wide format target spp abundance data
target_shore<-left_join(target_new, shorebird[,-c(2)], by = c("ZoneName", "SurveyYear"))%>%
  drop_na()
```

## Regression approach

Goal - assess relationships between visitation and year on target spp abundance
Approach - generalized additive model (spp abundance ~ year + visitation + zone)
Results - 

```{r Target spp gam approach}
### get grand means and join w/ plot data
target_90_home <- left_join(left_join(target_90_home, 
                            target_90_home %>% group_by(SpeciesCode)%>%
                            summarise(grand_mean = mean(Cover_mean)), 
                            by = c('SpeciesCode')),
                    # join with bird data
                  select(shorebird, -ZoneClass)) %>%
                  # first word of scientific name
                  mutate(Scientific_name = case_when(
                         Scientific_name == 'Mussel' ~ 'Mytilus',
                         TRUE ~ word(Scientific_name)),
                  # remove panel letters from zones (i.e. (A))
                  ZoneName = substr(ZoneName, 5, nchar(ZoneName)))

### write loop to run GAM across sites + target species

# make output df for GAM results
gam_output <- tibble(ZoneName = target_90_home$ZoneName,
                     Scientific_name = target_90_home$Scientific_name) %>%
              distinct() %>%
              mutate(`Adjusted R²` = NA,
                     `Deviance explained` = NA,
                      N = NA,
                     `Intercept - P` = NA,
                     `Year - F` = NA,
                     `Year - P` = NA,
                     `People - F` = NA,
                     `People - P` = NA)

# nested loop (i is for row # in gam_output table)
for (i in 1:nrow(gam_output)) {

    # subset relevant data 
    gam_data <- target_90_home %>%
    filter(Scientific_name == gam_output$Scientific_name[i] & 
           ZoneName == gam_output$ZoneName[i])
    
    # run model
    gam_mod <- gam(Cover_mean ~ s(SurveyYear) + s(people), data = gam_data)
    
    # put relevant summary info in output table 
    gam_output$N[i] <- nrow(gam_data)
    gam_output$`Deviance explained`[i] <- summary(gam_mod)$dev.expl
    gam_output$`Adjusted R²`[i] <- summary(gam_mod)$r.sq
    gam_output$`Intercept - P`[i] <- summary(gam_mod)$p.table[,4]
    gam_output$`Year - F`[i] <- summary(gam_mod)$s.table[1,3]
    gam_output$`Year - P`[i] <- summary(gam_mod)$s.table[1,4]
    gam_output$`People - F`[i] <- summary(gam_mod)$s.table[2,3]
    gam_output$`People - P`[i] <- summary(gam_mod)$s.table[2,4]
                
    # plot results (2 panels - one for people and one for year)
    p1 <- ggplot(data = gam_data,
           mapping = aes(x = SurveyYear, y = Cover_mean)) + 
            geom_point(color = 'black') + 
            xlab('Year') + 
            ylab('Percent cover') + 
            ggtitle(paste('GAM Result for', gam_data$Scientific_name[i], 'in',
                          gam_data$ZoneName[i]), 
              subtitle = paste('Abundance over time (P = ', scales::pvalue(gam_output$`Year - P`[i]),
                             ', F = ', round(gam_output$`Year - F`[i], digits = 2), 
                             ', N = ', gam_output$N, ')', sep = '')) +
            lltheme_light
    
    # if gam is significant, add in line
    if(gam_output$`Year - P`[i] <= 0.05) { 
      p1 <- p1 + geom_smooth(method = 'loess') }
    
    
    p2 <- ggplot(data = gam_data,
           mapping = aes(x = people, y = Cover_mean)) + 
            geom_point(color = 'black') + 
            xlab('Mean visitor count') + 
            ylab('Percent cover') + 
            ggtitle(paste('GAM Result for', gam_data$Scientific_name[i]
                          , 'in', gam_data$ZoneName[i]), 
              subtitle = paste('Abundance as a function of visitation (P = ', scales::pvalue(gam_output$`People - P`[i]),
                             ', F = ', round(gam_output$`People - F`[i], digits = 2), 
                             ', N = ', gam_output$N, ')', sep = '')) +
            lltheme_light
    
      # add line if GAM significant for people vs cover
      if(gam_output$`People - P`[i] <= 0.05) { 
      p2 <- p2 + geom_smooth(method = 'loess') }

    p3 <- p1 + p2
      
    ggsave(filename = paste('figures/GAM/gam_cover_people_time_', 
                 gsub('/', '', gam_data$Scientific_name[i]), '_',
                 gsub(' ', '', gam_data$ZoneName[i]), '.png', sep = ''),
           plot = p3,
           width = 12)
}

# clean up environment
remove(p1, p2, p3, site_list, spp_list, gam_data, gam_mod, i)

# make a formatted table using gam_output and the gt package
gt::gtsave(gam_output %>%
  mutate(`Intercept - P` = scales::pvalue(`Intercept - P`),
         `Year - P` = scales::pvalue(`Year - P`),
         `People - P` = scales::pvalue(`People - P`),
         `Adjusted R²` = round(`Adjusted R²`, digits = 2),
         `Deviance explained` = round(`Deviance explained`, digits = 2),
         `Year - F` = round(`Year - F`, digits = 2),
         `People - F` = round(`People - F`, digits = 2)) %>%
  rename(Site = ZoneName, `Target spp.` = Scientific_name) %>%
  gt::gt() %>%
    gt::fmt_percent(columns = `Deviance explained`, decimals = 0) %>%
    gt::cols_align(align = 'center', columns = `Intercept - P`:`People - P`),
  'figures/GAM/gam_table.png')

# add GAM info to dataset and make paneled plots for spp over time
target_90_home2 <- left_join(target_90_home, gam_output, 
                             by = c("ZoneName", "Scientific_name")) %>%
# add common names for figs 
  mutate(CommonName = case_when(
    Scientific_name == 'Balanus/Chthamalus' ~ 'Acorn barnacle',
    Scientific_name == 'Mytilus' ~ 'Mussel', 
    Scientific_name == 'Silvetia' ~ 'Golden rockweed',
    Scientific_name == 'Pollicipes' ~ 'Goose barnacle'
  )) %>%
# add numbers for factor order in paneled figs
  mutate(Fct_num = case_when (
    Scientific_name == 'Balanus/Chthamalus' ~ 1,
    Scientific_name == 'Mytilus' ~ 2, 
    Scientific_name == 'Silvetia' ~ 4,
    Scientific_name == 'Pollicipes' ~ 3
  )) %>%
  arrange(Fct_num)

target_90_home2$CommonName <- factor(target_90_home2$CommonName, 
                                     levels = c('Acorn barnacle',
                                                'Mussel', 
                                                'Goose barnacle',
                                                'Golden rockweed'))


# make paneled plot (panels are spp)  
# cover as a function of survey year

gam_time_fig <- function(dataset, savename) {

ggplot(data = dataset,
       mapping = aes(x = SurveyYear, y = Cover_mean, color = CommonName, group = ZoneName)) + 
  geom_point(size = 2) +
  # only show smooth lines for significant relationships from GAM
  geom_smooth(data = filter(dataset, `Year - P` <= 0.05),
              mapping = aes(x = SurveyYear, y = Cover_mean,
                            color = CommonName, fill = CommonName),
              method = 'loess') +
  # put info (r2 and pval) for each in color in text
  geom_text(data = select(ungroup(dataset), ZoneName, CommonName, 
                    `Deviance explained`, `Year - P`) %>% distinct() %>%
               mutate(Value = case_when(ZoneName == "Zone I" ~ 100,
                                        ZoneName == "Zone II" ~ 93,
                                        ZoneName == "Zone III" ~ 86)), 
             mapping = aes(x = 1991, y = Value, 
                           label = paste('P =', scales::pvalue(`Year - P`))),
             hjust = 0) +
  #scale_color_viridis(discrete = TRUE, begin = 0.4, end = 1, name = 'Site') +
  #scale_fill_viridis(discrete = TRUE, begin = 0.4, end = 1, name = 'Site') +
  xlab('Year') + 
  ylab('Percent cover') +
  facet_wrap(~CommonName, scales = 'free_y') + 
  coord_cartesian(xlim = c(1990, 2020), ylim = c(0,100)) + 
  scale_x_continuous(breaks = c(1990, 1995, 2000, 2005, 2010, 2015, 2020))  +
  lltheme_dark +
  theme(legend.position = 'none')
  # if using scientific names for strip/facet text
  # + theme(strip.text = element_text(face = "italic"))
  
  ggsave(paste('figures/GAM/all_panels_commonname', savename, '.png', sep = ''), width = 12)
  }

# make for Z1 only

gam_time_fig(filter(target_90_home2, ZoneName == 'Zone I'), 'time2')

# now make for people 
ggplot(data = filter(target_90_home2, ZoneName == 'Zone I'),
       mapping = aes(x = people, y = Cover_mean, color = CommonName)) + 
  geom_point(size = 2) +
  # only show smooth lines for significant relationships from GAM
  geom_smooth(data = filter(target_90_home2, ZoneName == 'Zone I' & `People - P` <= 0.05),
              mapping = aes(x = people, y = Cover_mean,
                            color = CommonName, fill = CommonName),
              method = 'loess') +
  # put info (r2 and pval) for each in color in text
  geom_text(data = select(ungroup(filter(target_90_home2, ZoneName == 'Zone I')), 
                          CommonName, `Deviance explained`, `People - P`) %>%
              distinct(),
            mapping = aes(x = 0, y = 100, 
                          label = paste('P =', scales::pvalue(`People - P`))),
            hjust = 0) +
  #scale_color_viridis(discrete = TRUE, begin = 0.4, end = 1, name = 'Site') +
  #scale_fill_viridis(discrete = TRUE, begin = 0.4, end = 1, name = 'Site') +
  xlab('Number of visitors') + 
  ylab('Percent cover') +
  facet_wrap(~CommonName) + 
  coord_cartesian(ylim = c(0,100)) + 
  lltheme_dark + 
  theme(legend.position = 'none')
  #theme(strip.text = element_text(face = "italic"))
  
  ggsave(paste('figures/GAM/all_panels_commonname2_people.png'), width = 12)



# all zones (color by spp) like LS
ggplot(data = target_90_home,
       mapping = aes(x = SurveyYear, y = Cover_mean, color = CommonName)) + 
  geom_point(size = 2) +
  # only show smooth lines for significant relationships from GAM
  geom_smooth(data = filter(dataset, `Year - P` <= 0.05),
              mapping = aes(x = SurveyYear, y = Cover_mean,
                            color = ZoneName, fill = ZoneName),
              method = 'loess') +
  # put info (r2 and pval) for each in color in text
  geom_text(data = select(ungroup(dataset), ZoneName, Scientific_name, 
                   `Deviance explained`, `Year - P`) %>% distinct() %>%
              mutate(Value = case_when(ZoneName == "Zone I" ~ 100,
                                       ZoneName == "Zone II" ~ 93,
                                       ZoneName == "Zone III" ~ 86)), 
            mapping = aes(x = 1991, y = Value, 
                          label = paste('P =', scales::pvalue(`Year - P`))),
            hjust = 0) +
  xlab('Year') + 
  ylab('Percent cover') +
  facet_wrap(~CommonName) + 
  coord_cartesian(xlim = c(1990, 2020), ylim = c(0,100)) + 
  scale_x_continuous(breaks = c(1990, 1995, 2000, 2005, 2010, 2015, 2020))  +
  lltheme_dark
  
  ggsave(paste('figures/GAM/all_panels_', savename, '.png', sep = ''), width = 12)


# plot people x cover figure
ggplot(data = target_90_home,
       mapping = aes(x = people, y = Cover_mean, color = ZoneName)) + 
  geom_point(size = 2) +
  # only show smooth lines for significant relationships from GAM
  geom_smooth(data = filter(target_90_home, `People - P` <= 0.05),
              mapping = aes(x = people, y = Cover_mean,
                            color = ZoneName, fill = ZoneName),
              method = 'loess') +
  # put info (r2 and pval) for each in color in text
  geom_text(data = select(ungroup(target_90_home), ZoneName, Scientific_name, 
                   `Deviance explained`, `People - P`) %>% distinct() %>%
              mutate(Value = case_when(ZoneName == "Zone I" ~ 100,
                                       ZoneName == "Zone II" ~ 93,
                                       ZoneName == "Zone III" ~ 86)), 
            mapping = aes(x = 0, y = Value, 
                          label = paste('P =', scales::pvalue(`People - P`))),
            hjust = 0) +
  scale_color_viridis(discrete = TRUE, begin = 0.4, end = 1, name = 'Site') +
  scale_fill_viridis(discrete = TRUE, begin = 0.4, end = 1, name = 'Site') +
  xlab('Number of visitors') + 
  ylab('Percent cover') +
  facet_wrap(~Scientific_name) + 
  coord_cartesian(ylim = c(0,100)) + 
  lltheme_dark + 
  theme(strip.text = element_text(face = "italic"))
  
  ggsave('figures/GAM/all_panels_people.png', width = 12)  
```


```{r}
#separate out people/year from target_shore
environ<-target_shore[c(1,2,3,4,5,17)] #tree

#separate out species matrix, including linking factors
species<-target_shore[c(1,2,3,8,9,10,11,12,13,14,15,16)] #bird
```

```{r}
#permanova- following rpubs.com/collnell/manova

#make matrix
species.matrix<-as.matrix(species[,4:12])

#standardize with sqrt
species.mat<-sqrt(species.matrix)

#find ecological resemblance
species.dist<-vegdist(species.mat, method='bray')

species.div<-adonis2(species.dist~ZoneName, data=species, permutations = 999, method='bray',strata = "PLOT")
species.div #issue with strata. ran w/out strata and pvalue=0.001

dispersion<-betadisper(species.dist, group=species$ZoneName)
permutest(dispersion) #pvalue=0.678

plot(dispersion, hull=F, ellipse=T) #just zone III shows up
# note from LP - i saw all zones, but it took a while to show up
```


```{r LP Suggestion PERMANOVA}
# let's run these parallel to the GAM analyses - running a separate model for each plot type
# does spp makeup of plots vary across zone and visitation?

# https://archetypalecology.wordpress.com/2018/02/21/permutational-multivariate-analysis-of-variance-permanova-in-r-preliminary/

# output from loop

# list to loop across (plot types)
spp_list <- unique(target_shore$Zone)

for(i in 1:length(spp_list)) {
  
  # filter for spp i
  dist_data <- target_shore %>%
    filter(Zone == spp_list[i])
  
  # get distance matrix
  dist_matrix <- vegdist(select(ungroup(dist_data), BARESUB:NOMATCH), method='bray')
  
  # conduct PERMANOVA (is makeup different across zones and visitation and year?)
  dist_mod <- adonis2(select(ungroup(dist_data), BARESUB:NOMATCH) ~ ZoneName + people + SurveyYear, data = dist_data)
  # different across zones, people and interaction (zones differ in visitation)
  
  # put results in table and save
  results <- tibble(Term = c('Zone', 'Visitation', 'Year', 'Residual', 'Total'),
                    `Deg. of freedom` = dist_mod$Df,
                    `Sum of squares` = round(dist_mod$SumOfSqs, digits = 2),
                    `F-value` = round(dist_mod$F, digits = 2),
                    `P-value` = dist_mod$`Pr(>F)`) %>%
            gt::gt() %>%
            gt::gtsave(paste('figures/PERMANOVA/permanova_result_',spp_list[i] ,'.png', sep = ''))

# add more things to this loop :)
  
}

```
#NMDS
#zone 3 NMDS
```{r}

color<-tibble(ZoneName=unique(species$ZoneName), Colors=c("green", "red", "blue"))

#make matrix
#9445 first row of zone3, 14028
#zone3.matrix<-as.matrix(select(species[which(species$ZoneName=="(C) Zone III")], -c("SiteCode","SiteName","ZoneName")))

zone3.matrix<-ungroup(species)%>% #ungroup b/c r wants to group first 3 columns
  filter(ZoneName=="(C) Zone III")%>% #filter by zone 3 for zone 3 nmds
  select(-c(1:3))%>% #take out non-num variables to be able to square
  as.matrix() #make matrix

#standardize with sqrt
zone3.mat<-sqrt(zone3.matrix)

#remove matrix rows with all 0s to run MDS? Error
#zone3.matrix=zone3.matrix[, colSums(count != 0) > 0]

zone3MDS<-metaMDS(zone3.mat, distance="bray", k=3, trymax = 35, autotransform = T) #stress=0.1252488
zone3MDS

stressplot(zone3MDS)

#graph doesn't work
NMDS1zone3<-zone3MDS$points[,1]
NMDS2zone3<-zone3MDS$points[,2]

zone3species<-species[which(species$ZoneName=="(C) Zone III"),]
zone3environ<-environ[which(environ$ZoneName=="(C) Zone III"),]
zone3species.plot<-cbind(zone3species, NMDS1zone3, NMDS2zone3, zone3environ)
#remove repeating columns
zone3species.plot<-zone3species.plot[-c(15:17)]
zone3species.plot<-left_join(zone3species.plot, color)


p<-ggplot(zone3species.plot, aes(NMDS1zone3, NMDS2zone3, color=Zone))+
  stat_ellipse(type='t',size = 1)+ ##draws 95% confidence interval ellipses
  theme_minimal()+
  annotate("text", x=min(NMDS1zone3), y=min(NMDS2zone3), label=paste('Stress =',round(zone3MDS$stress,3)))+ #add stress to plot
  geom_point(data=zone3species.plot, aes(NMDS1zone3, NMDS2zone3, shape=Zone))+
  scale_shape_manual(values=c(3,16,17,18))
p #make diff shapes for each Zone (species)
#polpol in zone 1 are similar to polpol in zone 2 and 3. pol and cht same across all zones. Myt and sil differs across zones more. 
#sil and myt diff in zone 1 than other zones

#fit mussel environ
mus.zone3 <- target_shore %>%
  filter(Zone == 'MYT'&ZoneName=="(C) Zone III") %>%
  select(-c(SiteCode:Zone, people))

mus.envzone3 <- target_shore %>%
  filter(Zone == 'MYT'&ZoneName=="(C) Zone III") %>%
  select(SiteName, SurveyYear, people)

musMDSzone3<-metaMDS(mus.zone3, distance= "bray", k = 2, trymax = 35, autotransform = T)
fitzone3<-envfit(musMDSzone3, mus.envzone3)
arrowzone3<-data.frame(fitzone3$vectors$arrows,R = fitzone3$vectors$r, P = fitzone3$vectors$pvals)
arrowzone3$FG <- rownames(arrowzone3)
arrow.pzone3<-filter(arrowzone3, P <= 0.05)


#graph with arrows as environmental variables
ggplot(zone3species.plot, aes(NMDS1zone3, NMDS2zone3))+
  #geom_point(data=zone3species.plot, aes(NMDS1zone3, NMDS2zone3, color=Zone),position=position_jitter(.1))+##separates overlapping points
  stat_ellipse(aes(fill=Zone), alpha=.2,type='t',size =1, geom="polygon")+
  theme_minimal()+
   geom_segment(data=arrow.pzone3, aes(x=0, y=0, xend=NMDS1, yend=NMDS2, label=FG, lty=FG), arrowzone3=arrow(length=unit(.2, "cm")*arrow.pzone3$R))+ ##add arrows (scaled by R-squared value)
  annotate("text", x=min(NMDS1zone3), y=min(NMDS2zone3), label=paste('Stress =',round(zone3MDS$stress,3)))+ #add stress to plot
  geom_point(data=zone3species.plot, aes(NMDS1zone3, NMDS2zone3, shape=Zone))+
  scale_shape_manual(values=c(3,16,17,18))
ggsave("NMDS_zone3.png", width=6, height=4)

#graph with ellipses as zone
ggplot(zone3species.plot, aes(NMDS1zone3, NMDS2zone3))+
  #geom_point(data=zone3species.plot, aes(NMDS1zone3, NMDS2zone3, color=Zone),position=position_jitter(.1))+##separates overlapping points
  stat_ellipse(aes(fill=ZoneName), alpha=.2,type='t',size =1, geom="polygon")+
  theme_minimal()+
   geom_segment(data=arrow.pzone3, aes(x=0, y=0, xend=NMDS1, yend=NMDS2, label=FG, lty=FG), arrowzone3=arrow(length=unit(.2, "cm")*arrow.pzone3$R))+ ##add arrows (scaled by R-squared value)
  annotate("text", x=min(NMDS1zone3), y=min(NMDS2zone3), label=paste('Stress =',round(zone3MDS$stress,3)))+ #add stress to plot
  geom_point(data=zone3species.plot, aes(NMDS1zone3, NMDS2zone3, shape=Zone))+
  scale_shape_manual(values=c(3,16,17,18))+
  scale_fill_manual(values=zone3species.plot$Colors)
ggsave("NMDS_zone3ellip.png", width=6, height=4)

```

#NMDS with zone 2 and 3
```{r}
#make matrix
zone23.matrix<-as.matrix(species[104:305,4:9])

#standardize with sqrt
zone23.mat<-sqrt(zone23.matrix)

zone23MDS<-metaMDS(zone23.mat, distance="bray", k=2, trymax = 35, autotransform = T) #stress=0.1252488
zone23MDS

stressplot(zone23MDS)

#graph doesn't work
NMDS1zone23<-zone23MDS$points[,1]
NMDS2zone23<-zone23MDS$points[,2]

zone23species<-species[which(species$ZoneName!="(A) Zone I"),]
zone23environ<-environ[which(species$ZoneName!="(A) Zone I"),]
zone23species.plot<-cbind(zone23species, NMDS1zone23, NMDS2zone23, zone23environ)
#remove repeating columns
zone23species.plot<-zone23species.plot[-c(15:17)]
zone23species.plot<-left_join(zone23species.plot, color)

ggplot(zone23species.plot, aes(NMDS1zone23, NMDS2zone23, color=Zone))+
  stat_ellipse(type='t',size = 1)+ ##draws 95% confidence interval ellipses
  theme_minimal()+
  annotate("text", x=min(NMDS1zone23), y=min(NMDS2zone23), label=paste('Stress =',round(zone23MDS$stress,3)))+ #add stress to plot
  geom_point(data=zone23species.plot, aes(NMDS1zone23, NMDS2zone23, shape=Zone))+
  scale_shape_manual(values=c(3,16,17,18))
#make diff shapes for each Zone (species)
#polpol in zone 1 are similar to polpol in zone 2 and 3. pol and cht same across all zones. Myt and sil differs across zones more. 
#sil and myt diff in zone 1 than other zones

#fit mussel environ
mus.zone23 <- target_shore %>%
  filter(Zone == 'MYT'& ZoneName!="(A) Zone I") %>%
  select(-c(SiteCode:Zone, people))

mus.envzone23 <- target_shore %>%
  filter(Zone == 'MYT'& ZoneName!="(A) Zone I") %>%
  select(SiteName, SurveyYear, people)

musMDSzone23<-metaMDS(mus.zone23, distance= "bray", k = 2, trymax = 35, autotransform = T)
fitzone23<-envfit(musMDSzone23, mus.envzone23)
arrowzone23<-data.frame(fitzone23$vectors$arrows,R = fitzone23$vectors$r, P = fitzone23$vectors$pvals)
arrowzone23$FG <- rownames(arrowzone23)
arrow.pzone23<-filter(arrowzone23, P <= 0.05)

#graph with arrows as environmental variables
ggplot(zone23species.plot, aes(NMDS1zone23, NMDS2zone23))+
  #geom_point(data=zone23species.plot, aes(NMDS1zone23, NMDS2zone23, color=Zone),position=position_jitter(.1))+##separates overlapping points
  stat_ellipse(aes(fill=Zone), alpha=.2,type='t',size =1, geom="polygon")+
  theme_minimal()+
   geom_segment(data=arrow.pzone23, aes(x=0, y=0, xend=NMDS1, yend=NMDS2, label=FG, lty=FG), arrowzone23=arrow(length=unit(.2, "cm")*arrow.pzone23$R))+ ##add arrows (scaled by R-squared value)
  annotate("text", x=min(NMDS1zone23), y=min(NMDS2zone23), label=paste('Stress =',round(zone23MDS$stress,3)))+ #add stress to plot
  geom_point(data=zone23species.plot, aes(NMDS1zone23, NMDS2zone23, shape=Zone))+
  scale_shape_manual(values=c(3,16,17,18))
ggsave("NMDS_zone23.png", width=6, height=4)

#graph with ellipses as zones
ggplot(zone23species.plot, aes(NMDS1zone23, NMDS2zone23))+
  #geom_point(data=zone23species.plot, aes(NMDS1zone23, NMDS2zone23, color=Zone),position=position_jitter(.1))+##separates overlapping points
  stat_ellipse(aes(fill=ZoneName), alpha=.2,type='t',size =1, geom="polygon")+
  theme_minimal()+
   geom_segment(data=arrow.pzone23, aes(x=0, y=0, xend=NMDS1, yend=NMDS2, label=FG, lty=FG), arrowzone23=arrow(length=unit(.2, "cm")*arrow.pzone23$R))+ ##add arrows (scaled by R-squared value)
  annotate("text", x=min(NMDS1zone23), y=min(NMDS2zone23), label=paste('Stress =',round(zone23MDS$stress,3)))+ #add stress to plot
  geom_point(data=zone23species.plot, aes(NMDS1zone23, NMDS2zone23, shape=Zone))+
  scale_shape_manual(values=c(3,16,17,18))+
  scale_fill_manual(values=unique(zone23species.plot$Colors))
ggsave("NMDS_zone23ellip.png", width=6, height=4)
```

#All zones NMDS
```{r}
color<-tibble(ZoneName=unique(species$ZoneName), Colors=c("green", "red", "blue"))

#remove repeating columns
zone3species.plot<-zone3species.plot[-c(15:17)]
zone3species.plot<-left_join(zone3species.plot, color)
speciesMDS<-metaMDS(species.mat, distance="bray", k=2, trymax = 35, autotransform = T) #stress=0.1572664
speciesMDS

stressplot(speciesMDS)

#graph doesn't work
NMDS1<-speciesMDS$points[,1]
NMDS2<-speciesMDS$points[,2]
species.plot<-cbind(species, NMDS1, NMDS2, environ)
species.plot<-species.plot[-c(15:17)]
species.plot<-left_join(species.plot, color)

p<-ggplot(species.plot, aes(NMDS1, NMDS2, color=ZoneName))+
  stat_ellipse(type='t',size = 1)+ ##draws 95% confidence interval ellipses
  theme_minimal()+
  #geom_text(data=species.plot, aes(NMDS1, NMDS2, label=Zone))+ #position=position_jitter(.35))+
  annotate("text", x=min(NMDS1), y=min(NMDS2), label=paste('Stress =',round(speciesMDS$stress,3)))+ #add stress to plot
  geom_point(data=species.plot, aes(NMDS1, NMDS2, shape=Zone))+
  scale_shape_manual(values=c(3,16,17,18))
p #make diff shapes for each Zone (species)
#polpol in zone 1 are similar to polpol in zone 2 and 3. pol and cht same across all zones. Myt and sil differs across zones more. 
#sil and myt diff in zone 1 than other zones

#fit environmental variables
#fit<-envfit(speciesMDS, species.mat)
#arrow<-data.frame(fit$vectors$arrows,R = fit$vectors$r, P = fit$vectors$pvals)
#arrow$FG <- rownames(arrowsil)
#arrow.p<-filter(arrowsil, P <= 0.05)

#fit mussel environ
mus.species <- target_shore %>%
  filter(Zone == 'MYT') %>%
  select(-c(SiteCode:Zone, people))

mus.env <- target_shore %>%
  filter(Zone == 'MYT') %>%
  select(SiteName, SurveyYear, people)

musMDS<-metaMDS(mus.species, distance= "bray", k = 2, trymax = 35, autotransform = T)
fit<-envfit(musMDS, mus.env)
arrow<-data.frame(fit$vectors$arrows,R = fit$vectors$r, P = fit$vectors$pvals)
arrow$FG <- rownames(arrow)
arrow.p<-filter(arrow, P <= 0.05)

ggplot(data=species.plot, aes(NMDS1, NMDS2))+
  geom_point(data=species.plot, aes(NMDS1, NMDS2, color=ZoneName),position=position_jitter(.1))+##separates overlapping points
  stat_ellipse(aes(fill=ZoneName), alpha=.2,type='t',size =1, geom="polygon")+ ##changes shading on ellipses
  theme_minimal()+
  geom_segment(data=arrow.p, aes(x=0, y=0, xend=NMDS1, yend=NMDS2, label=FG, lty=FG), arrow=arrow(length=unit(.2, "cm")*arrow.p$R)) ##add arrows (scaled by R-squared value)

#graph with arrows of how species differ
ggplot(species.plot, aes(NMDS1, NMDS2))+
  #geom_point(data=species.plot, aes(NMDS1, NMDS2, color=ZoneName),position=position_jitter(.1))+##separates overlapping points
  stat_ellipse(aes(fill=Zone), alpha=.2,type='t',size =1, geom="polygon")+
  theme_minimal()+
   geom_segment(data=arrow.p, aes(x=0, y=0, xend=NMDS1, yend=NMDS2, label=FG, lty=FG), arrow=arrow(length=unit(.2, "cm")*arrow.p$R))+ ##add arrows (scaled by R-squared value)
  annotate("text", x=min(NMDS1), y=min(NMDS2), label=paste('Stress =',round(speciesMDS$stress,3)))+ #add stress to plot
  geom_point(data=species.plot, aes(NMDS1, NMDS2, shape=Zone))+
  scale_shape_manual(values=c(3,16,17,18))

#graph with arrows as environmental variables
ggplot(species.plot, aes(NMDS1, NMDS2))+
  #geom_point(data=species.plot, aes(NMDS1, NMDS2, color=ZoneName),position=position_jitter(.1))+##separates overlapping points
  stat_ellipse(aes(fill=ZoneName), alpha=.2,type='t',size =1, geom="polygon")+
  theme_minimal()+
   geom_segment(data=arrow.p, aes(x=0, y=0, xend=NMDS1, yend=NMDS2, label=FG, lty=FG), arrow=arrow(length=unit(.2, "cm")*arrow.p$R))+ ##add arrows (scaled by R-squared value)
  annotate("text", x=min(NMDS1), y=min(NMDS2), label=paste('Stress =',round(speciesMDS$stress,3)))+ #add stress to plot
  geom_point(data=species.plot, aes(NMDS1, NMDS2, shape=Zone))+
  scale_shape_manual(values=c(3,16,17,18))+
  scale_fill_manual(values=unique(species.plot$Colors))
ggsave("NMDS_allzones.png", width=6, height=4)

#R values sig for both people and survey year.
#people arrow driving x axis, associated with mussels
#Survey year driving y axis, not really associated with specific species. 
```

#plot zone 1 first. Then add zone 2 and 3. Or start zone 3 (no people), add zone 2 and 1 (show diff because of people on mussel communities)
